\begin{thebibliography}{3}
\newcommand{\enquote}[1]{``#1''}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\providecommand{\urlprefix}{URL }
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi
\providecommand{\eprint}[2][]{\url{#2}}



\bibitem[{Allaire \emph{et~al.}(2017)Allaire, {R Foundation}, Wickham, {Journal
  of Statistical Software}, Xie, Vaidyanathan, {Association for Computing
  Machinery}, Boettiger, {Elsevier}, Broman, Mueller, Quast, Pruim, Marwick,
  Wickham, Keyes, and Yu}]{CRAN:rticles}
Allaire J, {R Foundation}, Wickham H, {Journal of Statistical Software}, Xie Y,
  Vaidyanathan R, {Association for Computing Machinery}, Boettiger C,
  {Elsevier}, Broman K, Mueller K, Quast B, Pruim R, Marwick B, Wickham C,
  Keyes O, Yu M (2017).
\newblock \emph{rticles: Article Formats for R Markdown}.
\newblock R package version 0.4.1,
  \urlprefix\url{https://CRAN.R-project.org/package=rticles}.

\bibitem[{MacFarlane(2017)}]{pandoc}
MacFarlane J (2017).
\newblock \emph{Pandoc: A Universal Document Converter}.
\newblock Version 1.19.2.1, \urlprefix\url{http://pandoc.org}.

\bibitem[{Xie(2017)}]{CRAN:knitr}
Xie Y (2017).
\newblock \emph{knitr: A General-Purpose Package for Dynamic Report Generation
  in R}.
\newblock R package version 1.17, \urlprefix\url{https://yihui.name/knitr/}.

\bibitem[{Karl~W. Broman(2015)}]{b}
Karl~W. Broman (2015).
\newblock R/qtlcharts: interactive graphics for quantitative trait locus
  mapping.
\newblock {\emph Genetics}, 199:359--361, \urlprefix\url{http://www.genetics.org/content/genetics/199/2/359.full.pdf}.

\bibitem[{Dheeru Dua and Casey Graff(2017)}]{Dua:2019}
Dheeru Dua and Casey Graff (2017).
\newblock {UCI} machine learning repository, \urlprefix\url{https://archive.ics.uci.edu/ml/datasets/abalone}.

\bibitem[{Dirk Eddelbuettel and James~Joseph Balamuta(August 2017)}]{PeerJ:Rcpp}
Dirk Eddelbuettel and James~Joseph Balamuta (August 2017).
\newblock Extending \textit{R} with \textit{C++}: A brief introduction to
  \textit{Rcpp}.
\newblock {\emph PeerJ Preprints}, 5:e3188v1, \urlprefix \url{https://doi.org/10.7287/peerj.preprints.3188v1}.

\bibitem[{Max Kuhn(2020)}]{d}
Max Kuhn (2020).
\newblock {\emph caret: Classification and Regression Training}.
\newblock R package version 6.0-86, \urlprefix \url{https://CRAN.R-project.org/package=caret}.

\bibitem[{Daniel Lüdecke(2020)}]{e}
Daniel Lüdecke (2020).
\newblock {\emph sjPlot: Data Visualization for Statistics in Social Science}.
\newblock R package version 2.8.6, \urlprefix \url{https://CRAN.R-project.org/package=sjPlot}.

\bibitem[{Warwick \emph{et~al.}(1994) Nash, T.L. Sellers, S.R. Talbot, A.J. Cawthorn, and W.B. Ford}]{article}
Warwick Nash, T.L. Sellers, S.R. Talbot, A.J. Cawthorn, and W.B. Ford (1994).
\newblock 7he population biology of abalone (haliotis species) in tasmania.
  i. blacklip abalone (h. rubra) from the north coast and islands of bass
  strait.
\newblock {\emph Sea Fisheries Division, Technical Report No}, 48, \urlprefix \url{https://www.researchgate.net/publication/287546509_7he_Population_Biology_of_Abalone_Haliotis_species_in_Tasmania_I_Blacklip_Abalone_H_rubra_from_the_North_Coast_and_Islands_of_Bass_Strait}

\bibitem[{Barret Schloerke, Di~Cook, Joseph Larmarange, Francois Briatte, Moritz Marbach,
  Edwin Thoen, Amos Elberg, and Jason Crowley (2020)}]{c}
Barret Schloerke, Di~Cook, Joseph Larmarange, Francois Briatte, Moritz Marbach,
  Edwin Thoen, Amos Elberg, and Jason Crowley(2020).
\newblock {\emph GGally: Extension to 'ggplot2'}.
\newblock R package version 2.0.0, \urlprefix \url{https://CRAN.R-project.org/package=GGally}.

\bibitem[{Taiyun Wei and Viliam Simko(2017)}]{corrplot2017}
Taiyun Wei and Viliam Simko (2017).
\newblock {\emph R package "corrplot": Visualization of a Correlation Matrix}.
\newblock (Version 0.84), \urlprefix \url{https://github.com/taiyun/corrplot}.

\bibitem[{Hadley Wickham, Mara Averick, Jennifer Bryan, Winston Chang, Lucy~D'Agostino
  McGowan, Romain François, Garrett Grolemund, Alex Hayes, Lionel Henry, Jim
  Hester, Max Kuhn, Thomas~Lin Pedersen, Evan Miller, Stephan~Milton Bache,
  Kirill Müller, Jeroen Ooms, David Robinson, Dana~Paige Seidel, Vitalie
  Spinu, Kohske Takahashi, Davis Vaughan, Claus Wilke, Kara Woo, and Hiroaki
  Yutani(2019).}]{a}
Hadley Wickham, Mara Averick, Jennifer Bryan, Winston Chang, Lucy~D'Agostino
  McGowan, Romain François, Garrett Grolemund, Alex Hayes, Lionel Henry, Jim
  Hester, Max Kuhn, Thomas~Lin Pedersen, Evan Miller, Stephan~Milton Bache,
  Kirill Müller, Jeroen Ooms, David Robinson, Dana~Paige Seidel, Vitalie
  Spinu, Kohske Takahashi, Davis Vaughan, Claus Wilke, Kara Woo, and Hiroaki
  Yutani (2019).
\newblock Welcome to the {tidyverse}.
\newblock {\emph Journal of Open Source Software}, 4(43):1686.

\end{thebibliography}
