---
title: An Abalone-Age Investigation

# Use letters 'a', 'b', ... or digits '1', '2', ... for different
# affiliations and addresses. Use '~' or '""' to suppress the
# display of the affiliation and address.
author:
  - name: "| 470408957 | 480423142 | 490209370 | 490384806 | 490443251 |"   

# Optional: line of arbitrary text with additional information.
# Could be used, for example, to mention the bibliographic info in a post-print.
# If not specified, defaults to "This version was compiled on \today"
date_subtitle: Data2002 Group Project | November 2020

# For footer text  TODO(fold into template, allow free form two-authors)

# Place DOI URL or CRAN Package URL here
doi_footer: https://github.sydney.edu.au/jary8982/M09B_early_5.git

# Abstract
abstract: |
  In this report, we investigate whether the age of _Haliotis_ _Rubra_ _(Blacklip_ _Abalone)_ can be estimated from external physical attributes. We constructed and evaluated two multiple linear regression models using the Akaike Information Criterion (AIC). After refinement of the selected model, we found that given two weights, three dimensions, and the sexual maturity of an ablone, we could explain 62.8% of the the variance in our target variable. Provided these measurements, predictions could in turn be untransformed to generate age estimates for abalone.

# Optional: One or more keywords
#keywords:
#  - pdf
#  - vignette

# Paper size for the document, values of letter and a4
papersize: letter

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 9pt

# Optional: Force one-column layout, default is two-column
#one_column: true

# Optional: Enables lineo mode, but only if one_column mode is also true
#lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
numbersections: true

# Optional: Specify the depth of section number, default is 5
secnumdepth: 3

# Optional: Bibliography 
#bibliography: pinp

# Optional: include-after
#include-after: somefile.tex

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: false

# Customize footer, eg by referencing the vignette
footer_contents: "Data2002 Group Project"

# Optional: Enable a 'Draft' watermark on the document
#watermark: true

# Produce a pinp document
# 
# Other options that can be set here, shown with their defaults
#   keep_tex: TRUE
#   collapse: FALSE
#   citation_package: natbib
#
# Any other arguments to rmarkdown::pdf_document could be set 
# here as well
#
output:
  pinp::pinp:
    includes:
      after_body: bib.tex

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{Introduction to pinp}
  %\VignetteKeywords{pinp,vignette}
  %\VignettePackage{pinp}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r setup, include=FALSE, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE)
# install.packages("qtlcharts")
# install.packages("corrplot")
# install.packages("GGally")
# install.packages("ggfortify")
# install.packages("caret")
# install.packages("sjPlot")
# install.packages("ggcorrplot")
# install.packages("QuantPsyc")
library(QuantPsyc)
library(tidyverse)
library(qtlcharts)
library(corrplot)
library(GGally)
library(ggfortify)
library(caret)
library(sjPlot)
library(ggcorrplot)
library(corrplot)

cols = c("sex",	
             "length",
             "diam",
             "height",
             "weight_whole",
             "weight_shucked",
             "weight_viscera",
             "weight_shell",
             "rings")

raw_data = readr::read_csv("../Data/abalone.data", cols)

unfiltered = raw_data %>% mutate(sex = factor(sex, c("M", "F", "I")),
                       rings = as.integer(rings))

data = raw_data %>% mutate(sex = factor(sex, c("M", "F", "I")),
                       rings = as.integer(rings)) %>%
filter(height < 0.5)


```
# Introduction 
Marine biologists and conservationists often study the age and growth patterns of a species in order to understand its demographics in and across various ecosystems.
As a sought after commodity within the fishing industry, this is especially true of Abalone.
However, the classical method for determining an abalone's age is arduous and time inefficient; counting the rings in a specially prepared shell under a microscope (\cite{Dua:2019}).
We therefore aim to find a technique for estimating an abalone's age using only physical attributes which are easily and quickly measured. We will construct a multiple regression model in order to predict the number of rings an abalone has, and evaluate whether this model can effectively predict observed values and would therefore have any utility when applied to new observations.

# Data Set
This data pertains to _Haliotis_ _Rubra_, an Australian species of abalone found predominantly in cold waters, such as off the coast of Tasmania. 
The relevant data were originally collected by _the Marine Resources Division in Taroona, Tasmania_ to explore neural network techniques for estimating the age of abalone. The data were made available by the University of California Irvine Machine Learning Repository(\cite{Dua:2019}). The dataset contains 4177 observations upon 9 different variables, and it contains no missing values. Each variable describes some physical property - a weight, dimension, sex, ring count - of the observed abalone.

## 2.1 Variables  
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
\textbf{Name} & \textbf{Type} & \textbf{Description} \\
\hline

Sex & Factor & Male, female or infant \\
\hline
Length (mm) & Continuous & Longest shell measurement \\
\hline
Diameter (mm) & Continuous & Perpendicular to length \\
\hline
Height (mm) & Continuous & With meat in shell \\
\hline
Whole Weight (g) & Continuous & Whole abalone \\
\hline
Shucked Weight (g) & Continuous & Weight of meat \\
\hline
Viscera Weight (g) & Continuous & Gut weight (after bleeding) \\
\hline
Shell Weight (g) & Continuous & After being dried \\
\hline
Rings & Integer & Number of rings. +1.5 gives age in years \\
\hline
\end{tabular}
\end{center}

    
\newpage
## 2.2 Outliers  

Initial data exploration reveals two clear anomalies in the height variable. These two observations are so far from the range of all other 4175 observed values that they are considered to be erroneous, and are discarded from the dataset.

```{r, echo=FALSE, fig.width=3, fig.height=1, cache=TRUE}
ggplot(unfiltered) + aes(x = height) + geom_boxplot(color = "#00AFBB") + theme(text=element_text(size=6), panel.background = element_blank(), panel.grid.major = element_line(colour="white"), panel.grid.minor = element_line(colour="white"), panel.border = element_rect(colour="steelblue", fill=NA, size=1))
```

# Analysis 

## 3.1 Transformations

Prior to selecting an appropriate model, we must acknowledge that the observed variables do not demonstrate a linear relationship with the observed number of rings _(Appendix 1)_, and we cannot consider sexual maturity in its current state as a ternary factor. Due to the nature of the observed curves, the ideal transformations for the predictor variables were as follows: Log transformations for length, diameter, and all the weights, and a square root transformation to height.  Transforming the predicted variable (rings) to the square root of its logarithm proved ideal. Each predictor variable now adopts a linear relationship with the predictive variable _(Appendix 2)_, allowing for a linear regressive model to work appropriately. Additionally, the sexual maturity factor was encoded using a contrast matrix.

```{r, echo=FALSE, include = FALSE}
res = model.matrix(~sex, data)

numericised_sex = select(data, -sex)
numericised_sex$sex_f = res[,2]
numericised_sex$sex_i = res[,3]

transformed = numericised_sex %>% mutate(log_whole = log(weight_whole),
                log_shucked = log(weight_shucked),
                log_viscera = log(weight_viscera),
                log_shell = log(weight_shell),
                log_diam = log(diam),
                log_length = log(length),
                sqrt_height = sqrt(height),
                sqrt_log_rings = sqrt(log(rings))) %>%
              select(log_whole,
                log_shucked,
                log_viscera,
                log_shell,
                log_diam,
                log_length,
                sqrt_height,
                sex_f,
                sex_i,
                sqrt_log_rings)
```

## 3.2 Model Selection

Having conducted our transformations, models could now be constructed. Two models were constructed; one using forward stepping variable selection, and the other using backward stepping variable selection. Both of these models were evaluated considering their $R^2$ and AIC values. The produced models were remarkably similar in regard to the above criteria, and the only notable difference between them is the omission of diameter and length from the forward model. The produced models are shown in the table below. 

```{r, echo= FALSE, include=FALSE, warning = FALSE, fig.width=10, fig.height=5, cache=TRUE}
library(magrittr)
library(knitr)
M1 = lm(sqrt_log_rings ~ ., transformed)
# M1
# drop1(M1, test = "F")

step.back.aic = step(M1, direction = "backward", trace = FALSE)
# summary(step.back.aic)

M2 = lm(sqrt_log_rings ~ 1, transformed)

step.fwd.aic = step(M2, scope=list(lower=M2, upper=M1), direction = "forward",trace = FALSE)

# summary(step.fwd.aic)

sjPlot::tab_model(step.fwd.aic, step.back.aic, show.ci = FALSE,show.aic = TRUE,dv.labels = c("Forward model","Backward model"), CSS = list(css.tdata = "padding:2px;text-align:left;vertical-align:top;"))
# cat(tab$page.style)
```

\begin{center}
\begin{tabular}{c c c c c}
    & Forward Model & & Backward Model & \\
    \hline
    Predictors & Estimates & p & Estimates & p\\
    \hline
    (Intercept) & 1.43 & $\textbf{<0.001}$ & 1.45 & $\textbf{<0.001}$\\
    log shell & 0.11 & $\textbf{<0.001}$ & 0.11 & $\textbf{<0.001}$\\
    log shucked & -0.19 & $\textbf{<0.001}$ & -0.19 & $\textbf{<0.001}$\\
    log whole & 0.19 & $\textbf{<0.001}$ & 0.20 & $\textbf{<0.001}$\\
    sex infant & -0.02 & $\textbf{<0.001}$ & -0.01 & $\textbf{<0.001}$\\
    log viscera & -0.03 & $\textbf{<0.001}$ & -0.02 & $\textbf{<0.001}$\\
    sqrt height & 0.13 & $\textbf{0.007}$ & 0.12 & $\textbf{0.012}$\\
    log diameter &  & & 0.07 & $\textbf{0.005}$\\
    log length &  & & -0.08 & $\textbf{0.005}$\\
    \hline
    Observations & 4175 &  & 1.45 & \\
    $R^2/R^2$ adjusted & 0.647 / 0.647 & & 0.648 / 0.647 & \\
    AIC & -10882.310 & & -10887.886 & \\
\end{tabular}
\end{center}

\newpage
## 3.3 Assumption Checking 

### Forward Residual vs Fitted/QQ Plot
```{r, warning = FALSE, echo = FALSE, fig.width=3, fig.height=1.7, cache=TRUE}
# autoplot(step.fwd.aic,which=1)
x = transformed
x = x %>% mutate(lfitted = step.fwd.aic$fitted.values, lresid = step.fwd.aic$residuals )
forward_plot1 = ggplot(x, aes(x = lfitted ,y = lresid)) + geom_jitter(size=0.1, alpha=0.5, color = "black", shape=".") + geom_smooth(size=0.5) +
  labs(x = "Fitted", y = "Residuals") + theme(text=element_text(size=6), panel.background = element_blank(), panel.grid.major = element_line(colour="white"), panel.grid.minor = element_line(colour="white"), panel.border = element_rect(colour="steelblue", fill=NA, size=1))

# qqnorm(step.fwd.aic$residuals)
# qqline(step.fwd.aic$residuals)
forward_plot2 = ggplot(step.fwd.aic, aes(sample = step.fwd.aic$residuals)) + stat_qq(alpha=0.5, size=0.5, shape=1, color = "black") + stat_qq_line(colour="blue") + theme(text=element_text(size=6), panel.background = element_blank(), panel.grid.major = element_line(colour="white"), panel.grid.minor = element_line(colour="white"), panel.border = element_rect(colour="steelblue", fill=NA, size=1)) + labs(x = "Theoretical", y = "Sample")

gridExtra::grid.arrange(forward_plot1, forward_plot2, ncol=2)
```

### Backward Residual vs Fitted/QQ Plot
```{r, warning = FALSE, echo = FALSE, fig.width=3, fig.height=1.7, cache=TRUE}
# autoplot(step.fwd.aic,which=1)
x = transformed
x = x %>% mutate(lfitted = step.back.aic$fitted.values, lresid = step.back.aic$residuals )
back_plot1 = ggplot(x, aes(x = lfitted ,y = lresid)) + geom_jitter(size=0.1, alpha=0.5, color = "black", shape=".") + geom_smooth(size=0.5) +
  labs(x = "Fitted", y = "Residuals") + theme(text=element_text(size=6), panel.background = element_blank(), panel.grid.major = element_line(colour="white"), panel.grid.minor = element_line(colour="white"), panel.border = element_rect(colour="steelblue", fill=NA, size=1))

# autoplot(step.fwd.aic,which=2)
back_plot2 = ggplot(step.back.aic, aes(sample = step.back.aic$residuals)) + stat_qq(alpha=0.5, size=0.5, shape=1, color = "black") + stat_qq_line(colour="blue") + theme(text=element_text(size=6), panel.background = element_blank(), panel.grid.major = element_line(colour="white"), panel.grid.minor = element_line(colour="white"), panel.border = element_rect(colour="steelblue", fill=NA, size=1)) + labs(x = "Theoretical", y = "Sample")
# qqnorm(step.back.aic$residuals)
# qqline(step.back.aic$residuals)

gridExtra::grid.arrange(back_plot1, back_plot2, ncol=2)
```


We must state and justify our assumptions - for both models - to validate any inferences made in our results.

- __Linearity__: The residual plot displays no obvious curvature for either model, thus the linearity assumption is satisfied.
- __Independence__: The data were collected across 5 different regions in the Tasman Sea \textit{(Appendix 3)}, with no systematic or intentional collection grouping. Granted these facts, there is no reason to believe there is any dependence between observations. Hence, independence can reasonably be assumed. 
- __Homoskedasticity__: For both models, the residuals do not appear to be fanning out or changing over the range of fitted values. Thus the constant error variance assumption is met.
- __Normality__: The normality assumption is at least approximately satisfied. For the QQ plot of each model, the points are reasonably close to the diagonal line. Regardless, the sample size is large enough to rely upon the central limit theorem. 



# Results 

Since the models constructed using the forward and backward approach share the same adjusted $R^2$, the RMSE and MAE were computed for each in order to determine and justify the better model. Graphs are shown below.

```{r, warning = FALSE}
set.seed(1)
cv_forward= train(
    sqrt_log_rings ~ . - log_diam - log_length - sex_f,
    transformed,
    method = "lm",
    trtControl = trainControl(
        method = "cv", number = 10,
        verboseIter = FALSE
    )
)
```

```{r, warning = FALSE}
cv_back= train(
    sqrt_log_rings ~ . -sex_f,
    transformed,
    method = "lm",
    trtControl = trainControl(
        method = "cv", number = 10,
        verboseIter = FALSE
    )
)
```

```{r, fig.width=3.25, fig.height=1.5, cache=TRUE}
results = resamples(list(forward = cv_forward,
                         back = cv_back))
plot1 = ggplot(results, metric = "RMSE") +  labs(y = "RMSE") + theme(text=element_text(size=6), panel.background = element_blank(), panel.grid.major = element_line(colour="white"), panel.grid.minor = element_line(colour="white"), panel.border = element_rect(colour="steelblue", fill=NA, size=1))

plot2 = ggplot(results, metric = "MAE") + labs(y = "MAE")+ theme(text=element_text(size=6), panel.background = element_blank(), panel.grid.major = element_line(colour="white"), panel.grid.minor = element_line(colour="white"), panel.border = element_rect(colour="steelblue", fill=NA, size=1))

gridExtra::grid.arrange(plot1, plot2, ncol=2)
```

It is evident that the backward model is the better model, as it has a lower Residual Mean Square Error and Mean Absolute Error. 
It is worth noting that the p-values for all the original variables are statistically significant, excepting one of the sexual maturity factors produced from the dummy coding contrast matrix. The p-value for `sex_f` was consistent with the null hypothesis that the __gender__ of the abalone is immaterial, while `sex_i` was significant, indicating that the sexual __maturity__ of the abalone was meaningful. \par
It must be conceded that there is apparent multicollinearity within the dataset _(Appendix 4)_; which may reduce the precision of the estimate coefficients and lessen the statistical power of the model. This multicollinearity is to be expected. Living organisms tend to grow physically as they age, with a rate that decreases over time. Naturally, our measured values display that trend.\par 
  
  It is a challenge to address multicollinearity within a data set where all the variables are significant. The optimal solution is to omit collinear variables which are already well represented by similar measurements. In our specific case, this was a number of the weight variables. The two more significant weight variables - according to the standardized regression coefficients - were shucked weight and whole weight;

```{r, include=FALSE}
a = lm.beta(step.back.aic)
a[c('log_whole', 'log_shucked', 'log_viscera', 'log_shell')]
a[c('log_diam', 'log_length', 'sqrt_height', 'sex_i')]
```  
\vspace{2mm}
\begin{tabular}{c c c c}
  log whole & log shucked & log viscera & log shell \\
  $1.4790451$ & $-1.5000279$ & $-0.1885621$ & $0.8269786$ \\
\hline
  log diam & log length & sqrt height & sex infant \\
  $0.19408468$ & $-0.19821088$ & $0.06175294$ & $-0.06326012$
\end{tabular}
\vspace{2mm}
Thus the viscera weight and shell weight were ignored in our final model; 

\begin{align*}
  \widehat{\sqrt{log(rings)}} &= 1.330 + 0.297 log(whole) -0.243 log(shucked)\\
  & + 0.153 log(diameter) -0.079 log(length)\\
  & + 0.205\sqrt{height}-0.013 Sex_{infant}
\end{align*}

Our model can predict the square root of the log of the number of rings with 62.8% explainable variance when using all the provided variables, making for a respectable regressive model.

# Discussion and Conclusion
We have constructed a model that will approximate an abalone's age from easily measured attributes - a useful tool when monitoring large marine ecosystems, where research time is far better spent collecting and analysing observations than counting rings.

## 5.1 Limitations
- Our data only pertains to _Haliotis Rubra_. The model does not account for species, and cannot claim to perform generally among Haliotes. Any conservational or environmental inferences are thus limited. 
- As noted above, there is high collinearity among the weight variables, and together this reduces the usefulness of each. It would perhaps be more profitable to forego one of these measurements in favour of another that would add more breadth to our profile of the abalone.
- The data were only collected from waters surrounding Tasmania _(Appendix 3)_. Although Blacklip Abalone __is__ prevalent in these waters, they are found in coastal waters reaching all the way from lower NSW to lower WA. This restricts the usefulness of the model, since it can only be used with confidence for Abalone in Tasmanian waters - a portion of a much larger population.


# Appendix

\textbf{Appendix 1: Correlation matrix of initial dataset}
```{r, warning = FALSE, fig.width=2.5, fig.height=2.5, cache=TRUE}
fn_data = function(data, mapping, ...) {
  p = ggplot(data = data, mapping = mapping) +
    geom_point(shape=".", size=0.001) +
    geom_smooth(method=lm, fill = "red", color="red", se = FALSE, size=0.3) 
  p
}
GGally::ggpairs(data, columns = 1:9, lower = list(continuous = fn_data),
    upper = list(continuous = wrap("cor", size = 1))) + theme_bw(base_size = 2)
```

\textbf{Appendix 2: Correlation matrix of transformed variables}
```{r, warning = FALSE, fig.width=2.5, fig.height=2.5, cache=TRUE}
fn_data = function(data, mapping, ...) {
  p = ggplot(data = transformed, mapping = mapping) +
    geom_point(shape=".", size=0.001) +
    geom_smooth(method=lm, fill = "red", color="red", se = FALSE, size=0.3) 
  p
}
GGally::ggpairs(transformed, columns = 1:10, lower = list(continuous = fn_data),
    upper = list(continuous = wrap("cor", size = 1))) + theme_bw(base_size = 2)
```

\textbf{Appendix 3: Data Collection Sites (\cite{article})}
\begin{figure}[H]
    \begin{center}
    \includegraphics[width=0.35\textwidth, height=2in]{Independence} 
    \end{center}
\end{figure}

\newpage

\textbf{Appendix 4: Correlation Matrix}
```{r, warning = FALSE, fig.width=4, fig.height=2.75, cache=TRUE}
corr <- round(cor(transformed), 2)
ggcorrplot(corr, method = "square") 
```
<!-- Acknowledgements, and References, are automatically typeset after this -->

