# Initial Data Exploration

```{r}
library(tidyverse)
library(reshape2)
library(gridExtra)
library(GGally)
library(qtlcharts)
library(ggfortify)
```

# Names found in `abalone.names` information file.
```{r}
cols = c("sex",	
             "length",
             "diam",
             "height",
             "weight_whole",
             "weight_shucked",
             "weight_viscera",
             "weight_shell",
             "rings")
```

# Read the data from `abalone.data` file.
```{r}
data_url = "https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data"
raw_data = readr::read_csv(data_url, cols)
raw_data = readr::read_csv("Data/abalone.data", cols)
```
# Correct the automatic variable typing
```{r}
data = raw_data %>% mutate(sex = factor(sex, c("M", "F", "I")),
                       rings = as.integer(rings))
data
```

# Coerce into a long format where each entry contains only one weight measurement
# This allows a density plot comparing each weight measurement.

```{r}
long = gather(data, "weight_type", "grams", 5:8)
long
```
#regress rings on the variable length
```{r}
#lm1 = lm(rings ~ length,data = data)

#data = data %>% mutate(resid = lm1$residuals,fitted = lm1$fitted.values)

#ggplot(data, aes(x = data$length, y = data$rings)) + geom_point() + theme_classic(base_size = 26) +   labs(x = "length", y = "rings") + geom_smooth(method = "lm", se = FALSE)

#p2 = ggplot(data,aes(x = length, y = resid)) + geom_point(size = 1) + theme_classic(base_size = 30) + labs(x = "length",y = "Residual") + geom_hline(yintercept = 0) 
#p2 + geom_smooth(method = "loess", se = FALSE)


#Non-linear -> do transiformations
p1 = ggplot(data,aes(x = length,y = rings)) + geom_point(size = 0.5) + scale_y_log10() +   theme_classic(base_size = 20) +labs(x = "length",y = "rings") + geom_smooth(method = "lm", se = FALSE)
p1

data = data %>% mutate(lrings = log(rings))
lm2 = lm(lrings ~ length, data = data)
lm2
data = data %>% mutate(lfitted = lm2$fitted.values,lresid = lm2$residuals)
p1 = ggplot(data, aes(x = length,y = lrings)) + geom_point(size = 0.5) + theme_classic(base_size = 20) +   labs(x = "length",y = "Log rings") + geom_smooth(method = "lm", se = FALSE)
p1

p2 = ggplot(data, aes(x = length,y = lresid)) + geom_point(size = 0.5) + theme_classic(base_size = 20) + labs(x = "length", y = "Residual") +   geom_hline(yintercept = 0)
p2

#Assumption 4:normality
data %>% ggplot() +   aes(sample = lresid) +   geom_qq(size = 2) + geom_qq_line()

autoplot(lm2, which = 1:2)
```
#regress rings on the variable diam
```{r}
p1 = ggplot(data,aes(x = diam,y = rings)) + geom_point(size = 0.5) + scale_y_log10() +   theme_classic(base_size = 20) +labs(x = "diam",y = "rings") + geom_smooth(method = "lm", se = FALSE)
p1

data = data %>% mutate(lrings = log(rings))
lm2 = lm(lrings ~ diam, data = data)
lm2
data = data %>% mutate(lfitted = lm2$fitted.values,lresid = lm2$residuals)
p1 = ggplot(data, aes(x = diam,y = lrings)) + geom_point(size = 0.5) + theme_classic(base_size = 20) +   labs(x = "diam",y = "Log rings") + geom_smooth(method = "lm", se = FALSE)
p1

p2 = ggplot(data, aes(x = diam,y = lresid)) + geom_point(size = 0.5) + theme_classic(base_size = 20) + labs(x = "diam", y = "Residual") +   geom_hline(yintercept = 0)
p2

#Assumption 4:normality
data %>% ggplot() +   aes(sample = lresid) +   geom_qq(size = 2) + geom_qq_line()

autoplot(lm2, which = 1:2)
```

#regress rings on the variable height
```{r}
p1 = ggplot(data,aes(x = height,y = rings)) + geom_point(size = 0.5) + scale_y_log10() +   theme_classic(base_size = 20) +labs(x = "height",y = "rings") + geom_smooth(method = "lm", se = FALSE)
p1

data = data %>% mutate(lrings = log(rings))
lm2 = lm(lrings ~ height, data = data)
lm2
data = data %>% mutate(lfitted = lm2$fitted.values,lresid = lm2$residuals)
p1 = ggplot(data, aes(x = height,y = lrings)) + geom_point(size = 0.5) + theme_classic(base_size = 20) +   labs(x = "height",y = "Log rings") + geom_smooth(method = "lm", se = FALSE)
p1

p2 = ggplot(data, aes(x = height,y = lresid)) + geom_point(size = 0.5) + theme_classic(base_size = 20) + labs(x = "height", y = "Residual") +   geom_hline(yintercept = 0)
p2

#Assumption 4:normality
data %>% ggplot() +   aes(sample = lresid) +   geom_qq(size = 2) + geom_qq_line()

autoplot(lm2, which = 1:2)
```

#regress rings on the variable weight_whole

```{r}
p1 = ggplot(data,aes(x = weight_whole,y = rings)) + geom_point(size = 0.5) + scale_y_log10() +   theme_classic(base_size = 20) +labs(x = "weight_whole",y = "rings") + geom_smooth(method = "lm", se = FALSE)
p1

data = data %>% mutate(lrings = log(rings))
lm2 = lm(lrings ~ weight_whole, data = data)
lm2
data = data %>% mutate(lfitted = lm2$fitted.values,lresid = lm2$residuals)
p1 = ggplot(data, aes(x = weight_whole,y = lrings)) + geom_point(size = 0.5) + theme_classic(base_size = 20) +   labs(x = "weight_whole",y = "Log rings") + geom_smooth(method = "lm", se = FALSE)
p1

p2 = ggplot(data, aes(x = weight_whole,y = lresid)) + geom_point(size = 0.5) + theme_classic(base_size = 20) + labs(x = "weight_whole", y = "Residual") +   geom_hline(yintercept = 0)
p2

#Assumption 4:normality
data %>% ggplot() +   aes(sample = lresid) +   geom_qq(size = 2) + geom_qq_line()

autoplot(lm2, which = 1:2)
```


```{r}
lm3 = lm(rings ~ length + diam + height+ weight_whole,data)
summary(lm3)$coefficients %>% round(4)
```

```{r,message='hide'}
GGally::ggpairs(data) + theme_bw(base_size = 1)
```

```{r}
#qtlcharts::iplotCorr(data)
```

# Density plots of different weight measurements, split by sex/maturity
```{r}
ggplot(long, aes(x=grams, fill=weight_type)) + geom_density(alpha=.5) +
    facet_wrap(nrow=3, ~sex)
```
# Create dataframes pertaining only to adult/infantile abalone
```{r}
adult = filter(data, sex != "I")
infant = filter(data, sex == "I")
```

# Plot diameter vs height for adults and infants
```{r}
p1 = ggplot(infant, aes(x=diam, y=height)) + geom_point() + labs(title="INFANTS") + ylim(0,0.4)
p2 = ggplot(adult, aes(x=diam, y=height)) + geom_point() + labs(title="MATURE") + ylim(0,0.4)
grid.arrange(p1, p2, nrow = 1)
```

# Test if the mean weight is different across male and female abalone.
```{r}
mWeight = filter(data, sex=="M")$weight_whole
fWeight = filter(data, sex=="F")$weight_whole

ggplot(adult, aes(y=weight_whole)) + geom_boxplot() + facet_wrap(~sex)

(t = t.test(mWeight, fWeight, var.equal=TRUE))
t$p.val
```